---
title: "EDA_Project"
format: html
---

```{r}
# Packges
library(tidyverse)
library(kableExtra)
library(nflverse)
library(nflreadr)

# data set
nfl_passes <- read_csv("https://raw.githubusercontent.com/36-SURE/2025/main/data/nfl_passes.csv")
```
- Question #1: Completion % and average time to throw with SLR or t-test

```{r}
model_1 <- lm(completion_percentage ~ avg_time, data = ds_1)
summary(model_1)

```
### Interpreting the coefficients

> P-value is not significant


```{r}

ggplot(model_1, aes(x = avg_time, y = completion_percentage)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(
    title = "Completion Percentage vs. Average Time to Throw",
    x = "Average Time to Throw (seconds)",
    y = "Completion Percentage (%)"
  ) +
  theme_minimal()

```

### Check Assumptions:




- Question #2: logistics regression

```{r}
model_2 <- glm(completion ~ offense_formation + routes_ran, 
             data = ds_2, family = binomial)
msummary(model_2)
```

### Checking for Model Conditions:

They are all categorical 

### Checking model significance


## Checking for Model significance
**Model: Multiple Logistic Regression with all Predictor Variables**

```{r, include= FALSE, echo=FALSE}
model_2 <- glm(completion ~ offense_formation + routes_ran, data = ds_2, 
             family = binomial)
msummary(model)
```



- Two data visualizations exploring the questions—both must be multivariate 
(i.e., involving 2+ variables) and in different formats



- One clustering analysis



- Conclusions for the hypotheses based on your EDA and data visualizations


