---
title: "EDA_Project"
format: html
---

```{r}
# Packages 
library(tidyverse)
library(kableExtra)
library(nflverse)
library(nflreadr)
library(caret)
library(rms)
library(CalibrationCurves)
# data set
nfl_passes <- read_csv("https://raw.githubusercontent.com/36-SURE/2025/main/data/nfl_passes.csv")
```
- Question #1: Completion % and average time to throw with SLR or t-test

```{r}
model_1 <- lm(completion_percentage ~ avg_time, data = ds_1)
summary(model_1)
```
### Interpreting the coefficients

> P-value is not significant


```{r}

ggplot(model_1, aes(x = avg_time, y = completion_percentage)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(
    title = "Completion Percentage vs. Average Time to Throw",
    x = "Average Time to Throw (seconds)",
    y = "Completion Percentage (%)"
  ) +
  theme_minimal()

```

### Check Assumptions:

> 


- Question #2: logistic regression


```{r}
nfl_passes2 <- nfl_passes |>
  select(complete_pass, 
         route_ran,
         offense_formation) |>
  # convert categorical variables to factors
  mutate(across(
    c(complete_pass,
    route_ran,
    offense_formation),
    as.factor
  )) |>
  # remove missing values
  drop_na()
```


### Checking for Model Conditions:

They are all categorical so the linearity condition is automatically met.


## First Model (Version 1): Predicting if Pass was Completed


### Checking for Model significance

**Model: Multiple Logistic Regression with all Predictor Variables**

```{r}
model_2 <- glm(complete_pass ~ offense_formation + route_ran, data = nfl_passes2, 
             family = binomial)
msummary(model_2)
```
### Count Distributions for our Predictors (in case of class imbalances)

```{r}
# get counts for simplification 
table(nfl_passes2$route_ran)
table(nfl_passes2$offense_formation)
```
### Grouping Predictors

- The code below generates our newly grouped predictors. Here is how we reorganized each variable:

```{r}
# grouping routes ran based on general characteristics
nfl_passes3 <- nfl_passes2 |>
  mutate(route_type = case_when(
    route_ran %in% c("SLANT", "FLAT", "SCREEN", "HITCH") ~ "Short",
    route_ran %in% c("IN", "OUT", "CROSS") ~ "Intermediate",
    route_ran %in% c("GO", "POST", "CORNER", "WHEEL") ~ "Deep",
    TRUE ~ "Other"
  ))

nfl_passes3 <- nfl_passes3 |>
 filter(offense_formation %in% c("SHOTGUN", "EMPTY", "SINGLEBACK", "I_FORM")) |>
  rename("formation_type" = offense_formation)

# dropped the category WILDCAT, JUMBO and PISTOL (maybe combine PISTOL and JUMBO??)
nfl_passes3 <- nfl_passes3 |>
  mutate(across(c(route_type, formation_type), factor))

## Getting the counts again
table(nfl_passes3$route_type)
table(nfl_passes3$formation_type)
```


## First Model (Version 2): Predicting whether the pass was completed or not

With our predictors now grouped into simpler, more balanced categories, we re-ran the logistic regression to predict whether the pass was completed or not.


```{r}
model_3 <- glm(complete_pass ~ formation_type + route_type, data = nfl_passes3, 
             family = binomial)
msummary(model_3)

```
### Final Model Interpretation

```{r}
# exponentiating the coeffiecients of variables that were found to be statistically
# significant

exp(-0.39877) # log of odds of the intercept. What does it mean for the intercept to be significant?
exp(0.29915) # log of odds for `formation_typeSINGLEBACK`
exp(0.99247) # log of odds for `route_typeIntermediate`
exp(1.43257) # log of odds for `route_typeOther`
exp(1.55478) # log of odds for `route_typeShort`
```

$$logit(\widehat{\pi}) = - 0.39877 + 0.14419 (formation_typeI_FORM) + 
0.02726(formation_typeSHOTGUN) + 0.29915(formation_typeSINGLEBACK) + 
0.99247(route_typeIntermediate) + 1.43257(route_typeOther) + 1.55478(route_typeShort) $$ 

> The coefficient for `formation_typeSINGLEBACK` is 0.29915 meaning that the likelihood 
(or the odds) of players completing a pass when performing the singleback formation increases by 0.29915 log of odds (ratio of success to failure) compared to when players perform the empty formation. 
> When we exponentiate this coefficient...

> The coefficient for `route_typeIntermediate` is 0.99247 meaning that the likelihood of players completing a pass when they run an intermediate route (which includes ) increases by 0.99247 log of odds compared to when they run a deep route (which includes ).
> When we exponentiate this coefficient...


> The coefficient for `route_typeOther` is 1.43257 indicating that the likelihood 
of players completing a pass when they run other type of routes (which include ) increases by 
1.43257 log of odds compared to when they run a deep route.
> When we exponentiate this coefficient...

> The coefficient for `route_typeShort` is 1.55478 indicating that the likelihood 
of players completing a pas when they run a short route (which includes ) increases
by 1.55478 log of odds compared to when they run a deep route.
> When we exponentiate this coefficient...

> We can conclude that all routes have a higher likelihood of completing a pass 
than deep routes (GO, POST, CORNER, and WHEEL). Also, the only formation that seems
to have a statisticall significant change in getting a pass completed is the 
singleback formation, which has a higher likelihood of completing a pass than the
empty formation.


## Evaluating the Performance of this Model

### Confusion Matrix

- To evaluate performance, we will use a confusion matrix.

```{r}
prediction <- predict(model_3,newdata=nfl_passes3,type="response") 
caret::confusionMatrix(data=as.numeric(prediction >0.5),reference=nfl_passes3$complete_pass)
```


```{r}
# checking the levels of complete_pass
table(nfl_passes3$complete_pass)

# THERE IS SOMETHING WRONG WITH THIS MATRIX-------------------------------------

# predict probabilities and classify
predicted_Classes <- predict(model_3, newdata = nfl_passes3,type = "response") >= 0.5
actual_Classes <- nfl_passes3$complete_pass


confusionMatrix(factor(predicted_Classes), factor(actual_Classes), positive = "TRUE")

# Ensure both are factors with the same levels
predicted_Classes <- factor(predicted_Classes, levels = c(FALSE,TRUE))
actual_Classes <- factor(actual_Classes, levels = c(FALSE, TRUE))

# confusion matrix
confusionMatrix(factor(predicted_Classes), factor(actual_Classes), positive = "TRUE")



#THIS (MANUAL) MATRIX WORKS-----------------------------------------------------

pred <- predict(model_3, newdata = nfl_passes3)
pred_y <- as.numeric(pred > 0)
true_y <- as.numeric(nfl_passes3$complete_pass == 1)
true_pos <- (true_y == 1) & (pred_y == 1)
true_neg <- (true_y == 0) & (pred_y == 0)
false_pos <- (true_y == 0) & (pred_y == 1)
false_neg <- (true_y == 1) & (pred_y ==0)
conf_mat <- matrix(c(sum(true_pos), sum(false_pos),
                     sum(false_neg), sum(true_neg)), 2, 2)

colnames(conf_mat) <- c(TRUE, FALSE)
rownames(conf_mat) <- c(TRUE, FALSE)
conf_mat
(conf_mat[1,1]+conf_mat[2,2])/(conf_mat[1,1]+conf_mat[2,2]+conf_mat[1,2]+conf_mat[2,1]) #accuracy
(conf_mat[1,1])/(conf_mat[1,1]+conf_mat[2,1]) # Checking Specificity
(conf_mat[2,2])/(conf_mat[2,2]+conf_mat[1,2]) # checking sensitivity

nfl_passes3 %>% count(complete_pass)
5463/(2687+5463) #
```


### Calibrate plot


```{r}
# has something to do woth bootstrap
set.seed(100)
```



- Two data visualizations exploring the questions—both must be multivariate 
(i.e., involving 2+ variables) and in different formats



- One clustering analysis

> maybe let's combine/compare this with top 10 players from the table

- Conclusions for the hypotheses based on your EDA and data visualizations


